### 2015-02-23 replyToReferee.txt
Implemented all suggestions of the referee except the first

We thank the referee for taking the time to carefully read the
manuscript and offer constructive feedback.  All of the
recommendations have been implemented, and can be identified in the
version of the new manuscript with changes in red. For each item in
the referee report, the relevant locations of the manuscript changes
are the following:

1) Page ??

2) Pages 8-9. Rather than add the suggested new figure as part (d) of
figure 1, we opted to add it as a new figure 2.

3) Page 16-17

i) Page 4

ii) Page 8

We hope that with these modifications, the manuscript is now ready for
publication.