### 2015-02-26 replyToReferee.txt
Now responding to referee's comment 1

We thank the referee for taking the time to carefully read the
manuscript and offer constructive feedback.

Regarding suggestion #1, discussing the formal mathematical
relationship between sensitivity of eigenmodes and transient growth
using concepts of pseudospectra:

We have added a paragraph discussing this subject to the end of
Section III.  However, we are reluctant to assert too strong a
connection to the mathematical theory of pseudospectra due to a
technical complication.  In the 'destabilizing perturbations' section
of Trefethen to which you refer, eq (13), the linear operator is given
a perturbation of norm bounded by some finite epsilon, causing the
eigenvalues to move by an amount that continuously goes to 0 as
epsilon -> 0. However, all evidence suggests that when shear is
changed from 0 to any finite value in the universal instability,
eigenvalues change discontinuously, with the change not going to 0
continuously as Ls/Ln -> infinity. In particular, the eigenvalues
discontinuously change from unstable to stable for a range of k_y, as
discussed in Refs. [25]-[26]. The reason for this discontinuity is
presumably the fact that the z domain must be infinite to resolve
eigenmode structures as shear goes to zero.  For any nonzero shear,
there is always some z for which the argument of J_0 is changed
significantly compared to the zero-shear case. Thus, to analyze the
role of magnetic shear in the universal instability using the formal
theory of pseudospectra, a level of mathematical care is required that
is beyond the scope of the present manuscript.

All of the other recommendations of the referee have been implemented,
and can be identified in the version of the new manuscript with
changes in red. For each item in the referee report, the relevant
locations of the manuscript changes are the following:

2) Pages 8-9. Rather than add the suggested new figure as part (d) of
figure 1, we opted to add it as a new figure 2.

3) Page 16-17

i) Page 4

ii) Page 8

We hope that with these modifications, the manuscript is now ready for
publication.